Integration of Multiple Constraints in ACG

### 2013-09-22 lenls/paper.tex
LENLS: Wrote Limits of Modularity

We start with the extraction constraints. \cite{pogodalla2012controlling}
describes three different constraints on extraction, all of them expressed
using the Constraint pattern (Figure~\ref{fig:single-cont}). We can take the
union of these three G-ACGs to get the G-ACG on Figure~\ref{fig:merged-conts}.
Next, we can incorporate knowledge from another paper,
\cite{pogodalla2007generalizing}, by splitting the $\mathcal{L}_{Syn}$ lexicon
and introducing a new intermediate signature
(Figure~\ref{fig:conts-elab}). The new signature, $Syntax$, deals purely with
syntax without any issues of operator scope and its functions have lower-order
types. We will finally add our contribution, the signatures $Neg$ and $Agr$
and the lexicons $\mathcal{L}_{Neg}$ and $\mathcal{L}_{Agr}$ which implement
the negation phenomenon we discussed before and some simple notion of
agreement, respectively.

@@

We notice that while the three constraints introduced in
\cite{pogodalla2012controlling} are expressed in terms of the $Syn$ signature,
our constraints $Agr$ and $Neg$ use a different interface, the $Syntax$
signature, to constrain the languages being defined. The $Cont_i$ constraints
are not actually constraining syntax itself, but the syntax-semantics
interface. This is in some cases a deliberate decision, since e.g. $Cont_1$ is
the constraint enforcing that quantifier scope must not escape outside of
tensed clauses. The $Syntax$ signature is not equipped to talk about operator
scope and so the constraint is expressed above the $Syn$ signature. On the
other hand, constraints $Cont_2$ and $Cont_3$ from
\cite{pogodalla2012controlling} are purely syntactic. However, we see it as a
perk of our approach that we can ignore that and import the constraint
wholesale without having to adapt them.

@@

As was said in Section~\ref{sec:constraints}, we will proceed by splitting the
atomic types $N$ and $NP$ into $N_{NEG{=}F}$ and $N_{NEG{=}T}$, and
$NP_{NEG{=}F}$ and $NP_{NEG{=}T}$, respectively. We will use these new types to
distinguish whether an $N$ or $NP$ phrase contains a negative determiner that
can pair up with a negated verb. Where before we had a single type, we will
now have possibly multiple to account for the different values of the
arguments' $NEG$ features. Here are some representative examples:

@@

We see that the negative determiner \emph{aucun} can only attach to phrases
which do not already contain a free negative determiner and that the resulting
phrase is marked as having a free negative determiner. On the other hand, the
determiner \emph{le} does not have any interaction and is still able to take
any $N$ phrase, preserving any free negative determiners inside. The type of
the negative particle \emph{ne} modifies a verb (in the example above,
$N_{ne^{tv}_i}$, a transitive verb) and produces a new verb that makes sure
that at least one of its arguments contains a negative determiner. Verbs are
not agnostic about the $NEG$ feature. By default, they specifically demand it
to be false, since a negative determiner cannot be coupled with a non-negated
verb. Finally, the types of the relativizer \emph{que} tell us two other
things: the $NP$ trace is considered as containing no negative determiners,
and a relative clause does not care about or alter the presence of free
negative determiners in the $N$ it modifies.

@@

\subsection{Discussion}

>>>>

We start with the extraction constraints. \cite{pogodalla2012controlling}
describes three different constraints on extraction, all of them expressed
using the Constraint pattern (Figure~\ref{fig:single-cont}). We can take the
union of these three G-ACGs to get the G-ACG on Figure~\ref{fig:merged-conts}.
Next, we can incorporate knowledge from \cite{pogodalla2007generalizing}, by
splitting the $\mathcal{L}_{Syn}$ lexicon and introducing a new intermediate
signature (Figure~\ref{fig:conts-elab}). The new signature, $Syntax$, deals
purely with syntax without any issues of operator scope and its functions have
lower-order types than those in $Syn$. We will finally add our contribution,
the signatures $Neg$ and $Agr$ and the lexicons $\mathcal{L}_{Neg}$ and
$\mathcal{L}_{Agr}$ which implement the constraint on negation we discussed
before and some simple notion of agreement, respectively.

@@

We notice that while the three constraints introduced in
\cite{pogodalla2012controlling} are expressed in terms of the $Syn$ signature,
our constraints $Agr$ and $Neg$ use a different interface, the $Syntax$
signature, to constrain the languages being defined. The $Cont_i$ constraints
are not actually constraining syntax itself, but the syntax-semantics
interface. This is in some cases a deliberate decision, since e.g. $Cont_1$ is
the constraint enforcing that quantifier scope must not escape outside of
tensed clauses. The $Syntax$ signature is not equipped to talk about operator
scope and so the constraint is expressed above the $Syn$ signature. On the
other hand, constraints $Cont_2$ and $Cont_3$ from
\cite{pogodalla2012controlling} are purely syntactic and therefore could be
more easily expressed in terms of $Syntax$ instead of $Syn$. However, we see
it as a perk of our approach that we can ignore that and import the
constraints wholesale without having to adapt them.

@@

As was said in Section~\ref{sec:constraints}, we will proceed by splitting the
atomic types $N$ and $NP$ into $N_{NEG{=}F}$ and $N_{NEG{=}T}$, and
$NP_{NEG{=}F}$ and $NP_{NEG{=}T}$, respectively. We will use these new types
to distinguish whether an $N$ or $NP$ phrase contains a negative determiner
that can pair up with a negated verb. Where before we had a single type, we
will now have possibly multiple types to account for the different values of
the arguments' $NEG$ features. Here are some representative examples:

@@

We see that the negative determiner \emph{aucun} can only attach to phrases
which do not already contain a free negative determiner and that the resulting
phrase is marked as having a free negative determiner. On the other hand, the
determiner \emph{le} does not have any interaction with the new $NEG$ feature
and is still able to take any $N$ phrase, preserving any free negative
determiners inside. The type of the negative particle \emph{ne} modifies a
verb (in the example above, $N_{ne^{tv}_i}$, a transitive verb) and produces a
new verb that makes sure that at least one of its arguments contains a
negative determiner. Verbs are not agnostic about the $NEG$ feature. By
default, they specifically demand it to be false, since a negative determiner
cannot be coupled with a non-negated verb. Finally, the types of the
relativizer \emph{que} tell us two other things: the $NP$ trace is considered
as containing no negative determiners, and a relative clause does not care
about or alter the presence of free negative determiners in the $N$ it
modifies.

@@

\subsection{Limits of Modularity}

We have decomposed our grammar into 9 signatures and 8 lexicons. Besides
making the grammar more readable and easily navigable, it also makes the
grammar more attractive from a computational point of view. The complexity of
parsing the highest abstract-most structures and verifying all the constraints
is decomposed into smaller subtasks. A string can be parsed into its
representation at the $Syntax$ level fairly easily (compared to parsing it at
the higher levels) since $Syntax$ employs types of lower orders than the more
abstract signatures. Finding the syntactic structure for a string is thus
feasible. We then proceed to verifying the syntactic constraints. Here we can
observe that the lexicons $\mathcal{L}_{Neg}$ and $\mathcal{L}_{Agr}$ are mere
relabelings and therefore parsing (i.e. constraint verification) is trivial
(i.e. decidable). Now in order to determine the meaning of our sentence, we
have to continue parsing by finding an antecedent in $Syn$. This involves
higher-order types used to encode the various scope-bearing
operators. Finally, we can opt in to verify the high-level syntax-semantics
constraints which employ dependent types. However, the lexicons in this case
resemble relabelings and a semi-decidable procedure could be used to solve the
problem. Having the grammar decomposed like this opens it up for computational
processing by letting us engage in different tactics at different stages
instead of having to parse a single hidden structure which enforces
grammaticality w.r.t. every constraint. G-ACGs share these parks also with
classical ACGs but they let use this composition in another dimension, which
facilitates the writing of multiple constraints.

However, it is important to keep in mind that the composability demonstrated
above is limited only to the composition of constraints. If we would like to
extend our language with, e.g., a new lexical category, then every signature
and every lexicon would have to be extended to cover this new category. It is
also important to point out that the constraint signatures that we have been
composing are not light-weight objects. They are usually defined by taking an
existing signature that describes the language at a relevant level and by
strengthening (usually all) the types of that signature to enforce the
constraint\footnote{This makes the expressivity of model-theoretic syntactic
  formalisms formalisms such as Interaction Grammars, in which adding the
  lexical items for \emph{ne} and \emph{aucun} along with enforcing the $Neg$
  constraint takes only two tree descriptions \cite{perrier2007french}, even
  more striking when compared to our approach.}. However, most of these type
alterations are mundane and we believe that this is fertile ground for
applying metagrammars.
